First steps in synthetic algebraic geometry

### 2017-03-14 snippets/properties-of-a1.tex
Nonreducedness of the affine line

\begin{prop}\label{prop:characterization-open-subfunctor}
Let~$X \in \Zar(S)$ be a Zariski sheaf. A subfunctor~$U
\hookrightarrow X$ is open if and only if any of the following conditions,
formulated in the internal language of~$\Zar(S)$, is
satisfied.
\begin{enumerate}
\item For all~$x \? X$ there is an ideal~$J \subseteq \affl$ such
that~$\affl/J$ is quasicoherent and such that~$x \in U$ if and only if~$1 \in
J$. In formal language:
\begin{multline*}
  \qquad\qquad\Zar(S) \models
  \forall x\?X\_
  \exists J \subseteq \affl\_ \\
  \speak{$J$ is an ideal} \wedge
  \speak{$\affl/J$ is quasicoherent} \wedge
  (x \in U \Leftrightarrow 1 \in J). \end{multline*}
\item For all~$x \? X$ the quotient~$\affl/K$ modulo the radical ideal
\[ K \defeq \{ f\?\affl \,|\, \speak{$f$ \inv} \Rightarrow (x \in U) \}
\subseteq \affl \]
is quasicoherent.
\end{enumerate}
\end{prop}

The displayed conditions are only meaningful in an intuitionistic context as
provided by the big Zariski topos. For instance, the first condition would be
trivially satisfied in classical logic, for any subfunctor~$U \hookrightarrow
X$: Classically, we have~$(x \in U) \vee (x \not\in U)$. If~$x \in U$, we can
pick~$J = (1)$, and if~$x \not\in U$, we can pick~$J = (0)$ (such that the
quotient~$\affl/J$ is isomorphic to~$\affl$, thus finitely presented and
therefore in particular quasicoherent).\footnote{Strictly speaking,
incompability with classical logical surfaces even earlier: in our internal
quasicoherence condition. The map~$E \otimes_\affl A \to [\Spec(A), E]$ which
the condition demands to be bijective has hardly any chance to be surjective if
the law of excluded middle is available to define maps~$\Spec(A) \to E$ by case
distinction.}

\begin{cor}\label{cor:sufficient-criterion-open-subfunctor}
Let~$X \in \Zar(S)$ be a Zariski sheaf. Let~$U \hookrightarrow X$ be
a subfunctor. If
\[
  \Zar(S) \models
  \forall x\?X\_
  \bigvee_{n \geq 0} \exists f_1,\ldots,f_n\?\affl\_
  (x \in U \Leftrightarrow \bigvee_i \speak{$f_i$ \inv}),
\]
then the subfunctor is open.
\end{cor}

\begin{proof}[Proof of Corollary~\ref{cor:sufficient-criterion-open-subfunctor}]
We show that the assumption implies the displayed condition of
Proposition~\ref{prop:characterization-open-subfunctor} in the internal
language. Given elements~$f_1,\ldots,f_n$ as in the assumption, we
construct the ideal~$J \defeq (f_1,\ldots,f_n) \subseteq \affl$. The
quotient~$\affl/J$ is indeed quasicoherent, being finitely presented, and
the statement that~$1 \in J$ is equivalent to one of
the~$f_i$ being invertible by locality of~$\affl$.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{prop:characterization-open-subfunctor}]

\end{proof}

>>>>

\begin{prop}The following properties of~$\affl$, as seen from the internal
point of view of~$\Zar(S)$, are consequences of~$\affl$ satisfying the
quasicoherence condition (and~$\affl$ not being the trivial ring):
\begin{enumerate}
\item $\affl$ is a field in the sense that
$\forall x\?\affl\_ \neg(x = 0) \Rightarrow \speak{$x$ \inv}$.
\item $\affl$ is not a reduced ring: $\neg \forall x\?\affl\_ (\bigvee_{n \geq
0} x^n = 0) \Rightarrow x = 0.$
\item Any function~$\affl \to \affl$ is given by a unique polynomial
in~$\affl[T]$.
\item Any monic polynomial~$p \in \affl[T]$ of degree at least one does \notnot have a
zero.
\end{enumerate}\end{prop}

\begin{proof}\begin{enumerate}
\item Let $x\?\affl$ such that~$\neg(x=0)$. We consider the quasicoherence
condition for the finitely presented~$\affl$-algebra~$A \defeq \affl/(x)$.
Since~$\Spec(A) \cong \brak{x=0} = \brak{\bot} = \emptyset$, the condition says
that the canonical homomorphism
\[ \affl/(x) \longrightarrow [\emptyset, \affl] \]
is an isomorphism. Since its codomain is the zero algebra, so is~$\affl/(x)$.
Therefore~$1 \in (x)$, that is,~$x$ is invertible.
\item Assume that~$\affl$ is reduced. Then the set~$\Delta \defeq \{
\varepsilon \in \affl \,|\, \varepsilon^2 = 0 \}$ is equal to~$\{ 0 \}$.
By the quasicoherence criterion applied to the finitely
presented~$\affl$-algebra~$A \defeq \affl[T]/(T^2)$, the canonical map
\[ \affl[T]/(T^2) \longrightarrow [\Spec(\affl[T]/(T^2)), \affl] \cong
  [\Delta, \affl] \cong \affl \]
is an isomorphism. It maps~$[T]$ to zero (the value of~$T$ at~$0 \in \Delta$).
Thus~$T \in (T^2)$ and therefore~$1 = 0$ in~$\affl$. This is a contradiction.
\item Immediate by considering the quasicoherence condition for the finitely
presented~$\affl$-algebra~$A \defeq \affl[T]$ and noticing that~$\Spec(A) \cong
\affl$.
\item Let~$p \in \affl[T]$ be a monic polynomial of degree at least one. Assume
that~$p$ doesn't have a zero in~$\affl$. Then the spectrum of~$A \defeq
\affl[T]/(p)$ is empty. The quasicoherence condition for~$A$ therefore implies
that~$\affl[T]/(p)$ is zero. This means that~$p$ is invertible in~$\affl[T]$.
A basic lemma in commutative algebra (whose standard proof is constructive)
then implies that with the exception of the constant term in~$p$, all
coefficients are nilpotent. This contradicts the assumption that~$p$ is monic
of degree at least one.\qedhere
\end{enumerate}\end{proof}
